---
title: "Assessing Quality of Nighttime Lights Data"
#author: "Robert Marty"
#date: "`r Sys.Date()`"
#output: rmarkdown::html_vignette
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
vignette: >
  %\VignetteEngine{knitr::knitr}
  %\VignetteIndexEntry{Assess Quality}
  %\VignetteEncoding{UTF-8}
---

```{r init, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  warning = FALSE, 
  message = FALSE,
  comment = "#>"
)
```

### Overview <a name="overview"></a>

The quality of nighttime lights data can be impacted by a number of factors, particularly cloud cover. To facilitate analysis using high quality data, Black Marble (1) marks the quality of each pixel and (2) in some cases, uses data from a previous date to fill the value---using a temporally-gap filled NTL value. 

This page illustrates how to examine the quality of nighttime lights data.

* [Setup](#setup)
* [Daily data](#daily)
  * [Nighttime lights](#daily-ntl)
  * [Quality flag](#daily-quality)
  * [Nighttime lights using good quality observations](#daily-goodq)
  * [Nighttime lights using good quality observations without gap filling](#daily-nogap)
* [Monthly/annual data](#ma)
  * [Nighttime lights](#ma-ntl)
  * [Number of observations](#ma-numobs)
  * [Quality](#ma-quality)
  * [Nighttime lights using good quality observations](#ma-ntl_gq)

### Setup <a name="setup"></a>

We first load packages and obtain a polygon for a region of interest; for this example, we use Switzerland. 

```{r, results='hide'}
library(blackmarbler)
library(geodata)
library(sf)
library(raster)
library(ggplot2)
library(dplyr)
library(exactextractr)
library(lubridate)
library(tidyr)
library(geodata)
library(knitr)

bearer <- "BEARER-TOKEN-HERE"
```

```{r bearer, include = FALSE}
#bearer <- read.csv("~/Desktop/bearer_bm.csv")$token
bearer <- read.csv("https://www.dropbox.com/scl/fi/pipze9nvak5qo7pedvwb4/bearer_bm.csv?rlkey=bkpv62s657c5w9jbchpg2vvr7&dl=1")$token
```

```{r}
roi_sf <- gadm(country = "CHE", level=0, path = tempdir()) |> st_as_sf()
```

### Daily Data <a name="daily"></a>

Below shows an example examining quality for daily data (`VNP46A2`).

#### Nighttime Lights <a name="daily-ntl"></a>

We download data for January 1st, 2023. When the `variable` parameter is not specified, `bm_raster` creates a raster using the `Gap_Filled_DNB_BRDF-Corrected_NTL` variable for daily data.

```{r, results='hide'}
ntl_r <- bm_raster(roi_sf = roi_sf,
                   product_id = "VNP46A2",
                   date = "2023-01-01",
                   bearer = bearer,
                   variable = "Gap_Filled_DNB_BRDF-Corrected_NTL")
```

<details>
<summary>Show code to produce map</summary>
```{r, ntl_daily_r_map, eval=FALSE}
#### Prep data
ntl_m_r <- ntl_r |> raster::mask(roi_sf) 

ntl_df <- rasterToPoints(ntl_m_r, spatial = TRUE) |> as.data.frame()
names(ntl_df) <- c("value", "x", "y")

## Distribution is skewed, so log
ntl_df$value_adj <- log(ntl_df$value+1)

##### Map 
ggplot() +
  geom_raster(data = ntl_df, 
              aes(x = x, y = y, 
                  fill = value_adj)) +
  scale_fill_gradient2(low = "black",
                       mid = "yellow",
                       high = "red",
                       midpoint = 4) +
  coord_quickmap() + 
  theme_void() +
  theme(plot.title = element_text(face = "bold", hjust = 0.5),
        legend.position = "none")
```
</details>

```{r, ntl_daily_r_map, echo=FALSE}
```

We notice that a number of observations are missing, that are poor quality and are not gap-filled. To understand the extent of missing date, we can use the following code to determine (1) the total number of pixels that cover Switzerland, (2) the total number of non-`NA` nighttime light pixels, and (3) the proportion of non-`NA` pixels.

```{r}
n_pixel <- function(values, coverage_fraction){
  length(values)
}

n_non_na_pixel <- function(values, coverage_fraction){
  sum(!is.na(values))
}

n_pixel_num        <- exact_extract(ntl_r, roi_sf, n_pixel)
n_non_na_pixel_num <- exact_extract(ntl_r, roi_sf, n_non_na_pixel)

print(n_pixel_num)
print(n_non_na_pixel_num)
print(n_non_na_pixel_num / n_pixel_num)
```

By default, the `bm_extract` function computes these values:

```{r, results='hide'}
ntl_df <- bm_extract(roi_sf = roi_sf,
                     product_id = "VNP46A2",
                   date = seq.Date(from = ymd("2023-01-01"),
                                   to = ymd("2023-01-10"),
                                   by = 1),
                   bearer = bearer,
                   variable = "Gap_Filled_DNB_BRDF-Corrected_NTL")

knitr::kable(ntl_df)
```

The below figure shows trends in average nighttime lights (left) and the proportion of the country with a value for nighttime lights (right). For some days, low number of pixels corresponds to low nighttime lights (eg, January 3 and 5th); however, for other days, low number of pixels corresponds to higher nighttime lights (eg, January 9 and 10). On January 3 and 5, missing pixels could have been over typically high-lit areas (eg, cities)---while on January 9 and 10, missing pixels could have been over typically lower-lit areas. 

<details>
<summary>Show code to produce figure</summary>
```{r, n_obs_figure, eval=FALSE}

ntl_df %>%
  dplyr::select(date, ntl_mean, prop_non_na_pixels) %>%
  pivot_longer(cols = -date) %>%
  
  ggplot(aes(x = date,
             y = value)) + 
  geom_line() +
  facet_wrap(~name,
             scales = "free")
```
</details>

```{r, n_obs_figure, echo=FALSE}
```

#### Quality <a name="daily-quality"></a>

For daily data, the quality values are:

* 0: High-quality, Persistent nighttime lights

* 1: High-quality, Ephemeral nighttime Lights

* 2: Poor-quality, Outlier, potential cloud contamination, or other issues

* 255: No retrieval, Fill value (masked out on ingestion)

We can map quality by using the `Mandatory_Quality_Flag` variable.

```{r, results='hide'}
quality_r <- bm_raster(roi_sf = roi_sf,
                       product_id = "VNP46A2",
                       date = "2023-01-01",
                       bearer = bearer,
                       variable = "Mandatory_Quality_Flag")
```

<details>
<summary>Show code to produce map</summary>
```{r, quality_daily_r_map, eval=FALSE}
#### Prep data
quality_r <- quality_r |> mask(roi_sf) 

quality_df <- rasterToPoints(quality_r, spatial = TRUE) |> as.data.frame()
names(quality_df) <- c("value", "x", "y")

quality_df <- quality_df %>%
  dplyr::mutate(value_str = case_when(
    value == 0 ~ "0: High-quality, persistent",
    value == 1 ~ "1: High-quality, ephemeral",
    value == 2 ~ "2: Poor-quality",
    value == 255 ~ "255: Fill value"
  ))

##### Map 
ggplot() +
  geom_raster(data = quality_df, 
              aes(x = x, y = y, 
                  fill = value_str)) +
  scale_fill_brewer(palette = "Spectral", direction = -1) + 
  labs(fill = "Quality") +
  coord_quickmap() + 
  theme_void() +
  theme(plot.title = element_text(face = "bold", hjust = 0.5))
```
</details>

```{r, quality_daily_r_map, echo=FALSE}
```

#### Nighttime lights for good quality observations <a name="daily-goodq"></a>

The `quality_flag_rm` parameter determines which pixels are set to `NA` based on the quality indicator. By default, only pixels with a value of `255` are filtered out. However, if we only want data for good quality pixels, we can adjust the `quality_flag_rm` parameter.

```{r, results='hide'}
ntl_good_qual_r <- bm_raster(roi_sf = roi_sf,
                             product_id = "VNP46A2", 
                             date = "2023-01-01",
                             bearer = bearer,
                             variable = "Gap_Filled_DNB_BRDF-Corrected_NTL",
                             quality_flag_rm = c(2, 255))
```

<details>
<summary>Show code to produce map</summary>
```{r, ntl_daily_good_qual_map, eval=FALSE}
#### Prep data
ntl_good_qual_r <- ntl_good_qual_r |> mask(roi_sf) 

ntl_good_qual_df <- rasterToPoints(ntl_good_qual_r, spatial = TRUE) |> as.data.frame()
names(ntl_good_qual_df) <- c("value", "x", "y")

## Distribution is skewed, so log
ntl_good_qual_df$value_adj <- log(ntl_good_qual_df$value+1)

##### Map 
ggplot() +
  geom_raster(data = ntl_good_qual_df, 
              aes(x = x, y = y, 
                  fill = value_adj)) +
  scale_fill_gradient2(low = "black",
                       mid = "yellow",
                       high = "red",
                       midpoint = 4) +
  coord_quickmap() + 
  theme_void() +
  theme(plot.title = element_text(face = "bold", hjust = 0.5),
        legend.position = "none")
```
</details>

```{r, ntl_daily_good_qual_map, echo=FALSE}
```

#### Nighttime lights for good quality observations, without gap filling <a name="daily-nogap"></a>

By default, the `bm_raster` function uses the `Gap_Filled_DNB_BRDF-Corrected_NTL` variable for daily data. Gap filling indicates that some poor quality pixels use data from a previous date; the `Latest_High_Quality_Retrieval` indicates the date the nighttime lights value came from.

```{r, results='hide'}
ntl_tmp_gap_r <- bm_raster(roi_sf = roi_sf,
                           product_id = "VNP46A2",
                           date = "2023-01-01",
                           bearer = bearer,
                           variable = "Latest_High_Quality_Retrieval")
```

<details>
<summary>Show code to produce map</summary>
```{r, ntl_tmp_gap_map, eval=FALSE}
#### Prep data
ntl_tmp_gap_r <- ntl_tmp_gap_r |> mask(roi_sf) 

ntl_tmp_gap_df <- rasterToPoints(ntl_tmp_gap_r, spatial = TRUE) |> as.data.frame()
names(ntl_tmp_gap_df) <- c("value", "x", "y")

##### Map 
ggplot() +
  geom_raster(data = ntl_tmp_gap_df, 
              aes(x = x, y = y, 
                  fill = value)) +
  scale_fill_distiller(palette = "Spectral") +
  coord_quickmap() + 
  theme_void() +
  labs(fill = "Temporal\nGap\n(Days)",
       title = "Temporal gap between date (Jan 1, 2023)\nand date of high quality pixel used") +
  theme(plot.title = element_text(face = "bold", hjust = 0.5))
```
</details>

```{r, ntl_tmp_gap_map, echo=FALSE}
```

Instead of using `Gap_Filled_DNB_BRDF-Corrected_NTL`, we could ignore gap filled observations---using the `DNB_BRDF-Corrected_NTL` variable. Here, we also remove poor quality pixels.

```{r, results='hide'}
ntl_r <- bm_raster(roi_sf = roi_sf,
                   product_id = "VNP46A2",
                   date = "2023-01-01",
                   bearer = bearer,
                   variable = "DNB_BRDF-Corrected_NTL",
                   quality_flag_rm = c(2, 255))
```

<details>
<summary>Show code to produce map</summary>
```{r, ntl_daily_nogap_r_map, eval=FALSE}
#### Prep data
ntl_r <- ntl_r |> mask(roi_sf) 

ntl_df <- rasterToPoints(ntl_r, spatial = TRUE) |> as.data.frame()
names(ntl_df) <- c("value", "x", "y")

## Distribution is skewed, so log
ntl_df$value_adj <- log(ntl_df$value+1)

##### Map 
ggplot() +
  geom_raster(data = ntl_df, 
              aes(x = x, y = y, 
                  fill = value_adj)) +
  scale_fill_gradient2(low = "black",
                       mid = "yellow",
                       high = "red",
                       midpoint = 4) +
  coord_quickmap() + 
  theme_void() +
  theme(plot.title = element_text(face = "bold", hjust = 0.5),
        legend.position = "none")
```
</details>

```{r, ntl_daily_nogap_r_map, echo=FALSE}
```

### Monthly/Annual Data <a name="ma"></a>

Below shows an example examining quality for monthly data (`VNP46A3`). The same approach can be used for annual data (`VNP46A4`); the variables are the same for both monthly and annual data.

#### Nighttime Lights <a name="ma-ntl"></a>

We download data for January 2023. When the `variable` parameter is not specified, `bm_raster` creates a raster using the `NearNadir_Composite_Snow_Free` variable for monthly and annual data---which is nighttime lights, removing effects from snow cover. 

```{r, results='hide'}
ntl_r <- bm_raster(roi_sf = roi_sf,
                   product_id = "VNP46A3", 
                   date = "2023-01-01",
                   bearer = bearer,
                   variable = "NearNadir_Composite_Snow_Free")
```

<details>
<summary>Show code to produce map</summary>
```{r, ntl_r_map, eval=FALSE}
#### Prep data
ntl_r <- ntl_r |> mask(roi_sf) 

ntl_df <- rasterToPoints(ntl_r, spatial = TRUE) |> as.data.frame()
names(ntl_df) <- c("value", "x", "y")

## Distribution is skewed, so log
ntl_df$value_adj <- log(ntl_df$value+1)

##### Map 
ggplot() +
  geom_raster(data = ntl_df, 
              aes(x = x, y = y, 
                  fill = value_adj)) +
  scale_fill_gradient2(low = "black",
                       mid = "yellow",
                       high = "red",
                       midpoint = 4) +
  coord_quickmap() + 
  theme_void() +
  theme(plot.title = element_text(face = "bold", hjust = 0.5),
        legend.position = "none")
```
</details>

```{r, ntl_r_map, echo=FALSE}
```

#### Number of Observations <a name="ma-numobs"></a>

Black Marble removes poor quality observations, such as pixels covered by clouds. To determine the number of observations used to generate nighttime light values for each pixel, we add `_Num` to the variable name.

```{r, results='hide'}
cf_r <- bm_raster(roi_sf = roi_sf,
                  product_id = "VNP46A3",
                  date = "2023-01-01",
                  bearer = bearer,
                  variable = "NearNadir_Composite_Snow_Free_Num")
```

<details>
<summary>Show code to produce map</summary>
```{r, cf_r_map, eval=FALSE}
#### Prep data
cf_r <- cf_r |> mask(roi_sf) 

cf_df <- rasterToPoints(cf_r, spatial = TRUE) |> as.data.frame()
names(cf_df) <- c("value", "x", "y")

cf_df$value <- cf_df$value %>% as.factor()

##### Map 
ggplot() +
  geom_raster(data = cf_df, 
              aes(x = x, y = y, 
                  fill = value)) +
  scale_fill_viridis_d() + 
  labs(fill = "Number of\nObservations") +
  coord_quickmap() + 
  theme_void() +
  theme(plot.title = element_text(face = "bold", hjust = 0.5))
```
</details>

```{r, cf_r_map, echo=FALSE}
```

#### Quality <a name="ma-quality"></a>

For monthly and annual data, the quality values are:

* 0: Good-quality, The number of observations used for the composite is larger than 3

* 1: Poor-quality, The number of observations used for the composite is less than or equal to 3

* 2: Gap filled NTL based on historical data

* 255: Fill value

We can map quality by adding `_Quality` to the variable name.

```{r, results='hide'}
quality_r <- bm_raster(roi_sf = roi_sf,
                       product_id = "VNP46A3",
                       date = "2023-01-01",
                       bearer = bearer,
                       variable = "NearNadir_Composite_Snow_Free_Quality")
```

<details>
<summary>Show code to produce map</summary>
```{r, quality_r_map, eval=FALSE}
#### Prep data
quality_r <- quality_r |> mask(roi_sf) 

quality_df <- rasterToPoints(quality_r, spatial = TRUE) |> as.data.frame()
names(quality_df) <- c("value", "x", "y")

quality_df <- quality_df %>%
  dplyr::mutate(value_str = case_when(
    value == 0 ~ "0: Good quality",
    value == 1 ~ "1: Poor quality",
    value == 2 ~ "2: Gap filled",
    value == 255 ~ "255: Fill value"
  ))

##### Map 
ggplot() +
  geom_raster(data = quality_df, 
              aes(x = x, y = y, 
                  fill = value_str)) +
  scale_fill_brewer(palette = "Spectral", direction = -1) + 
  labs(fill = "Quality") +
  coord_quickmap() + 
  theme_void() +
  theme(plot.title = element_text(face = "bold", hjust = 0.5))
```
</details>

```{r, quality_r_map, echo=FALSE}
```

#### Nighttime lights for good quality observations <a name="ma-ntl_gq"></a>

The `quality_flag_rm` parameter determines which pixels are set to `NA` based on the quality indicator. By default, only pixels with a value of `255` are filtered out. However, if we also want to remove poor quality pixels, we can adjust the `quality_flag_rm` parameter.

```{r, results='hide'}
ntl_good_qual_r <- bm_raster(roi_sf = roi_sf,
                             product_id = "VNP46A3", 
                             date = "2023-01-01",
                             bearer = bearer,
                             variable = "NearNadir_Composite_Snow_Free",
                             quality_flag_rm = c(1, 255))
```

<details>
<summary>Show code to produce map</summary>
```{r, ntl_good_qual_map, eval=FALSE}
#### Prep data
ntl_good_qual_r <- ntl_good_qual_r |> mask(roi_sf) 

ntl_good_qual_df <- rasterToPoints(ntl_good_qual_r, spatial = TRUE) |> as.data.frame()
names(ntl_good_qual_df) <- c("value", "x", "y")

## Distribution is skewed, so log
ntl_good_qual_df$value_adj <- log(ntl_good_qual_df$value+1)

##### Map 
ggplot() +
  geom_raster(data = ntl_good_qual_df, 
              aes(x = x, y = y, 
                  fill = value_adj)) +
  scale_fill_gradient2(low = "black",
                       mid = "yellow",
                       high = "red",
                       midpoint = 4) +
  coord_quickmap() + 
  theme_void() +
  theme(plot.title = element_text(face = "bold", hjust = 0.5),
        legend.position = "none")
```
</details>

```{r, ntl_good_qual_map, echo=FALSE}
```

